\documentclass[12pt,oneside]{../../sfsuthesis} 
 
\RequirePackage{standalone}
\usepackage[draft]{../../MAThesisOutputFormat}
\input{../../preamble.tex}
\usepackage[backend=biber,style=numeric]{biblatex}
\addbibresource{../../thesis.bib}

\begin{document}

\chapter{Chow Rings and Bergmann Fans}

\section{Chow Rings}
\todo{Introduce the Chow Ring of a matroid. From lattice of flats to quotient ring}
\todo{Use our example matroid and construct its Chow Ring}

\subsection{Properties of Chow Rings}
\todo{Use words like \emph{homogeneous polynomial}, \emph{graded ring}, etc\dots}

\subsection{The Degree Map}
\todo{How do I explain this?
    I guess I can at least say it's linear and sends terms of full degree to 1.
    Maybe I'll understand it this time around}

\section{Bergmann Fans}
\todo{Quick definition of a fan?}

\subsection{How to make a Bergmann Fan}

\todo{Show definition from Chow Ring to Bergmann Fan}
\todo{Work our small example into a fan}

\subsection{Properties of Bergmann Fans}

\todo{Figure out their important properties. They're unimodal, so we'll include that.
    Oh, they're balanced as well (and so tropical). What am I missing?}
\todo{Define star of a fan. Reference A-H-K; star of a Bergmann fan is again a Bergmann fan? Or something like that}

\section{Relationship with the Characteristic Polynomial}

\todo{Come up with a nice way of relating the reduced characteristic polynomial with our ring (and therefore fan)}
\todo{Define \(\alpha\) and \(\beta\). Here or in a subsection? Or should it be up when we introduce the ring itself?}

\subsection{How to show Log-Concavity}
\todo{Offer a self-contained proof that log-concavity is equivalent to showing that one particular relationship between \(\alpha\) and \(\beta\) found in A-H-K.
    They have the proof, but it requires digging through citations.
    Should be able to consolidate it.}

\end{document}